Effects of PCSK9 Inhibition With Evolocumab on Lipoprotein Metabolism in subjects with elevated Lp(a)

### 2019-10-04 manuscript.Rmd
add new affiliation
new figure for LPa-apo(a)

Xiang Zhang, PhD$^1$,
Lotte C.A. Stiekema, MD$^2$,
Erik S.G. Stroes, PhD$^2$,
Albert K. Groen, PhD$^1$

1. Department of Experimental Vascular Medicine, Amsterdam University Medical Center, University of Amsterdam, Meibergdreef 9, 1105 AZ Amsterdam, The Netherlands
2. Department of Vascular Medicine, Amsterdam University Medical Center, University of Amsterdam, Meibergdreef 9, 1105 AZ Amsterdam, The Netherlands

@@

![Figure 1 Evolocumab treatment for 16 weeks resulted in mean (95% credible interval) reduction in Lp(a) compared to placebo group: 24.5% (14.0% to 35.5%)](./Treatment_effect_lpa.png)

@@

Lp(a) is thought to be comprised of a LDL particle and a covalently bound apo(a). 
In the Nightingale metabolomics platform, three LDL particles, including large, medium and small LDL, were profiled. 
At baseline, we identified that on average 43.9% (95% credible interval [0.0102%, 89.3%]) large LDL, 40.9% ([0.00807%, 88.6%]) medium LDL particles and 50.5% ([5.39% 97.3%]) small LDL particles were with apo(a) (Figure 3). 
The correlations of LDL-apo(a) proportion between these three LDL particles were -0.4 (large and medium LDL), -0.62 (large and small LDL) and -0.42 (medium and small LDL).

![Figure 3 Posterior mean of LDL-apo(a) proportion in large, medium and small LDL particles. Each dot (in total 4000 dots) is a mean proportion of LDL particles with apolipoprotein(a) sampled from the posterior distribtuions. The marginal distribution of the posterior mean was on the diagnal. The pearson correlation coefficients were shown](./Lpa_LDL_baseline.png)

After treatment of evolocumab, on average 76.5% (41.5%, 100%) large LDL particles were with apo(a). 
In the pool of medium LDL particles, on average 69.2% (25.5%, 100%) of them were with apo(a). 
Regarding small LDL particles, on average 75.8% (39.3%, 100%) of them were with apo(a).

![Figure 4 Posterior mean of LDL-apo(a) proportion in large, medium and small LDL particles. Each dot (in total 4000 dots) is a mean proportion of LDL particles with apolipoprotein(a) sampled from the posterior distribtuions. The marginal distribution of the posterior mean was on the diagnal. The pearson correlation coefficients were shown](./Lpa_LDL_visit2.png)

>>>>

Xiang Zhang, PhD$^1,^2$,
Lotte C.A. Stiekema, MD$^2$,
Erik S.G. Stroes, PhD$^2$,
Albert K. Groen, PhD$^1$

1. Department of Experimental Vascular Medicine, Amsterdam University Medical Center, University of Amsterdam, Meibergdreef 9, 1105 AZ Amsterdam, The Netherlands
2. Human and Animal Physiology, Wageningen University, De Elst 1, 6708 WD Wageningen, The Netherlands
3. Department of Vascular Medicine, Amsterdam University Medical Center, University of Amsterdam, Meibergdreef 9, 1105 AZ Amsterdam, The Netherlands

@@

![Figure 1 Evolocumab treatment for 16 weeks resulted in mean (95% credible interval) reduction in Lp(a) compared to placebo group: 18.9% (7.51% to 29.2%)](./Treatment_effect_lpa.png)

@@

Lp(a) is thought to be comprised of a LDL particle and a covalently bound apo(a). 
In the Nightingale metabolomics platform, three LDL particles, including large, medium and small LDL, were profiled. 
At baseline, we identified that on average 56% (95% credible interval [9%, 100%]) large LDL, 50% ([3%, 97%]) medium LDL particles and 53% ([7% 99%]) small LDL particles were with apo(a) (Figure 3).

After treatment of evolocumab, on average 70% (28%, 100%) large LDL particles were with apo(a). 
In the pool of medium LDL particles, on average 61% (11%, 99%) of them were with apo(a). 
Regarding small LDL particles, on average 68% (24%, 100%) of them were with apo(a).

![Figure 3 Proportions of particles with and without apo(a) in large, medium and small LDL pools. The sizes of pie charts were scaled to the average particle concentrations.](./Lpa_LDL_composition.png)
